---
title: "subcell-correlation"
format: html
execute:
  freeze: true
  warning: false
---

# Alignment of adjacent slides

```{r}
#| label: libraries
#| include: false
#| echo: false
source("R/fct_correlation.R")
library(shinySTRegister)
library(SpatialExperiment)
library(data.table)
```

```{r}
#| label: constants

NBINS <- 200

```

::: panel-tabset
## CosMX

Of the three replicates (Lung5), two adjacent ones (2 and 3) were lined up manually (using `shinySTRegister`).

```{r}
lung5 <- cbind(
  readRDS("data/cosmx/aligned/Lung5_Rep2_spe.rds"),
  load_cosmx_as_spatial_experiment(
    "data/cosmx/Lung5_Rep3_exprMat_file.csv",
    "data/cosmx/Lung5_Rep3_metadata_file.csv"))
```

```{r}
binned_lung5 <- convert_spe_to_image(lung5)
```

```{r}
#| output-location: slide
plot_counts(binned_lung5)
```

### Spot-based correlation across transcripts

```{r}
sum_counts_per_bin(lung5, NBINS) |>
  plot_counts_correlations()
```

## Merscope

```{r}
ovarian2 <- cbind(
  readRDS("data/merscope/aligned/HumanOvarianCancerPatient2Slice1_spe.rds"),
  load_merscope_as_spatial_experiment(
    "data/merscope/HumanOvarianCancerPatient2Slice3_cell_by_gene.csv",
    "data/merscope/HumanOvarianCancerPatient2Slice3_cell_metadata.csv"))
```

```{r}
binned_ovarian2 <- convert_spe_to_image(ovarian2)
```

```{r}
#| output-location: slide
plot_counts(binned_ovarian2)
```

### Spot-based correlation across transcripts

```{r}
sum_counts_per_bin(ovarian2, NBINS) |>
  plot_counts_correlations()
```

## Xenium

```{r}
breast <- cbind(
  readRDS("data/xenium/aligned/Xenium_FFPE_Human_Breast_Cancer_Rep1_spe.rds"),
  load_xenium_as_spatial_experiment(
    "data/xenium/Xenium_FFPE_Human_Breast_Cancer_Rep2_cell_feature_matrix.csv",
    "data/xenium/Xenium_FFPE_Human_Breast_Cancer_Rep2_cells.csv"))
```

```{r}
binned_breast <- convert_spe_to_image(breast)
```

```{r}
#| output-location: slide
plot_counts(binned_breast)
```

### Spot-based correlation across transcripts

```{r}
sum_counts_per_bin(breast, NBINS) |>
  plot_counts_correlations()
```
:::

```{r}
#| include: false

saveRDS(list(CosMx = lung5,
             MERSCOPE = ovarian2,
             Xenium = breast),
        "data/aligned_samples_list.rds")

```

### RNiftyReg vs. manual alignment

These alignments were performed by hand, automated alignment (using `RNiftyReg`) failed in all cases except for CosMX. Even in this case, the results were slightly worse than with manual alignment.

```{r}
#| label: nifty_v_manual
#| echo: false

nifty_lung <- readRDS("data/lung5s23_niftyreg.rds")
comparison_cor_data <- data.frame(
  correlation = sum_counts_per_bin(nifty_lung, NBINS) |>
    get_binwise_correlation_from_counts(),
  type = "nifty"
) |>
  rbind(
    data.frame(
      correlation = sum_counts_per_bin(lung5, NBINS) |>
        get_binwise_correlation_from_counts(),
      type = "manual"
    )
  )

ggplot(comparison_cor_data, aes(x = correlation, color = type)) +
  geom_density()
```
