---
title: "Comparison"
execute:
  freeze: true
  warning: false
---

```{r}
#| label: libraries
#| include: false
#| echo: false
source("R/fct_correlation.R")
library(shinySTRegister)
library(SpatialExperiment)
library(data.table)
sample_list <- readRDS("data/aligned_samples_list.rds")
```


# Spot-based correlation across technologies
Spot-based correlation was calculated as
$$
cor(S_{i,j}, S'_{i,j})\,,\; i,j \in [1, 2, ..., 200]
$$

::: {.panel-tabset group="nbins"}

## 300 bins
```{r}
#| echo: false
get_correlation_data_frame(sample_list, nbins = 300) |>
ggplot(aes(x = Correlation, color = Technology)) +
  geom_density() +
  theme_minimal()
```
## 200 bins
```{r}
#| echo: false
get_correlation_data_frame(sample_list, nbins = 200) |>
ggplot(aes(x = Correlation, color = Technology)) +
  geom_density() +
  theme_minimal()
```

## 100 bins
```{r}
#| echo: false
get_correlation_data_frame(sample_list, nbins = 100) |>
ggplot(aes(x = Correlation, color = Technology)) +
  geom_density() +
  theme_minimal()
```
:::

# Feature-based correlation across technologies
Feature-based correlation was calculated as
$$
cor(f_{i}, f'_{i})\,,\; i \in [1, 2, ..., N_{features}]
$$

::: {.panel-tabset group="nbins"}

## 300 bins
```{r}
#| echo: false
get_correlation_data_frame(sample_list, "gene", nbins = 300) |>
ggplot(aes(x = Correlation, color = Technology)) +
  geom_density() +
  theme_minimal()
```

## 200 bins
```{r}
#| echo: false
get_correlation_data_frame(sample_list, "gene", nbins = 200) |>
ggplot(aes(x = Correlation, color = Technology)) +
  geom_density() +
  theme_minimal()
```

## 100 bins
```{r}
#| echo: false
get_correlation_data_frame(sample_list, "gene", nbins = 100) |>
ggplot(aes(x = Correlation, color = Technology)) +
  geom_density() +
  theme_minimal()
```
:::

## Note

These results show a clear difference in the technical variability of these technologies *as applied to different tissues*. Differences in tissue heterogeneity, genes measured and/or sample surface are likely to influence this comparison.
